Agreement Attraction in Turkish: Effects of Nominal and Verbal Plural Morphemes

### 2019-07-03 Paper/AgreementAttraction.Rmd
Some stuff. I don't remember.

Attraction errors in the production and comprehension of subject-verb agreement, in which a verb does not agree with the grammatical agreement controller, but with a potential attractor, have been the main topic of research in many studies for quite a long time. In fact, it is still a widely researched area in psycholinguistic studies. Despite the comprehensive research that has been carried out, studies that have been conducted on agreement attraction in Turkish have been very limited. In fact, @Lago is currently the only study that explores this phenomenon in Turkish. @Lago  makes use of genitive-possessive structures in the subject position, in which the possessive-marked noun is the head of the noun phrase which acts as the grammatical agreement controller, and the genitive noun serves as a potential attractor. In a speeded acceptability judgment study, @Lago found a significant effect of number agreement attraction. However, the interpretation of their findings may be a result of the fact that non-subjecthood cues originate from their use of morphologically ambiguous forms of the possessive. In the possessive forms that are used, all the head nouns in their stimuli are ambigiuous between possessive and accusative.

@@

df_lagoetal_unp <- df_lagoetal %>% 
                    subset(is.na(Grammatical)) %>%
                    dplyr::select(-Grammatical:-Label)
df_lagoetal_attr <- df_lagoetal %>% 
                  subset(!is.na(Grammatical)) %>%
                  dplyr::select(-Distance:-NewCond)

@@

@@

df_merge_exp1 <- data_clean %>% ungroup() %>% 
                      dplyr::select(source=experiment, grammatical, attractor_num, # condition,
                                    subject, item=Item,
                                    ResponseYes, RT)
df_merge_exp1$experiment <- "Experiment 1"
df_merge_exp1$grammatical <- with(df_merge_exp1, ifelse(grammatical == "gram", 
                                                        "grammatical",
                                                        "ungrammatical"))
df_merge_exp1$attractor_num <- with(df_merge_exp1, ifelse(attractor_num == "pl", 
                                                          "plural", 
                                                          "singular"))
df_merge_exp1$item %<>% as.integer()
df_merge_exp1$subject %<>% as.character()

df_lagoetal_attr %<>% mutate(ResponseYes = (Response == "yes") )
df_merge_lago <- df_lagoetal_attr %>% ungroup() %>% 
                      dplyr::select(grammatical=Grammatical, attractor_num=Attractor, #condition=Condition, 
                                    subject=Participant, item=Item,
                                    ResponseYes, RT)
df_merge_lago$experiment <- "Lago et al. (2018)" 
df_merge_lago$source <- NA
df_merge_lago$item %<>% add(1000)

@@

avg_clean <- df_merged %>%
                group_by(experiment, source, grammatical, attractor_num) %>%
                summarize(avgRT = mean(RT), p_yes = mean(ResponseYes, na.rm = T))
```

@@

```{r exp1ModelResponse, eval = F}

@@

df_merged %<>% within(., {
  cGrammatical <- ifelse(grammatical == "grammatical", .5, -.5)
  cUngrammatical <- ifelse(grammatical == "ungrammatical", .5, -.5)
  cAttractorPlural <- ifelse(attractor_num == "plural", .5, -.5)
  cEndsInVowel <- ifelse(experiment == "Experiment 1", .5, -.5)
})
df_merged_nofillers <- df_merged %>% subset(is.na(source) | source != "filler")

# ## test model parameterization using a simple GLM first
# m <- glm(ResponseYes ~ cEndsInVowel * cUngrammatical * cAttractorPlural,
#          data = df_merged_nofillers,
#          family = binomial("probit"))
# summary(m)

@@

m_responses <- brm(ResponseYes ~ cEndsInVowel * cUngrammatical * cAttractorPlural + 
                                 (cUngrammatical * cAttractorPlural + 1| subject) +
                                 (cUngrammatical * cAttractorPlural + 1| item),
                   data = df_merged_nofillers,
                   family = bernoulli("probit"),
                   file = fname_exp1_responses, 
                   chains = n_chains, cores = n_cores, iter = n_iter)

@@

```{r exp1ResponseModelPlot, eval = F, fig.height= 1.5, message=FALSE, warning=FALSE, fig.cap="Estimates and 95% credible intervals for the analysis of the probability of a 'yes' response."}

contrast_names <- c("cUngrammatical" = "Ungrammaticality",
                    "cAttractorMatch" = "Attactor Match",
                    "cUngrammatical:cAttractorMatch" = "Ungrammaticality * Attractor Match")

>>>>

Attraction attraction in the processing of subject-verb agreement have been an illuminating with regard to \ldots.

<!--
Introduction:
 - Review of the key attraction findings: 
        * English. 
      * German: @ReifegersteEtAl:2016
      * Definitely Russian (Sliosar on gender). [@SlioussarMalko:2016]
      * Arabic: @TuckerEtAl:2015
      * Spanish: @LagoEtAl:2015
      * Turkish: @LagoEtAl:2018
      * Hammerly 2018 [@HammerlyEtAl:2018]
      * @LinzenLeonard:2018
      * @SmithFranck:2018
      * @ParkerPhillips:2017
      * @VasishthEtAl:2017
      * @ReifegersteEtAl:2016
      * @KwonSturt:2016
      * @RisticMolinaroMancini:2016
      * @SlevcMartin:2016
      * @PatsonHusband:2015
      * @Enochson:2015
      * @FranckColonnaRizzi:2015
      * @WagersEtAl:2009
      * Slovak: @BadeckerKuminiak:2007
      * @BockEtAl:2001
      * @FiederEtAl:2014
      * @Staub:2009
      * @TannerNicolBrehm:2014
      * @XiangEtAl:2013
      * @NicenboimEtAl:2018

      @TannerBulkes:2015
      @Staub:2010
      @DankEtAl:2014
      @PearlmutterGarnseyBock:1999
      @FooteBock:2012
      @DillonEtAl:2013

  - Accounts of 
-->

errors in the production and comprehension of subject-verb agreement,

in which a verb does not agree with the grammatical agreement controller, but with a potential attractor, have been the main topic of research in many studies for quite a long time. In fact, it is still a widely researched area in psycholinguistic studies. Despite the comprehensive research that has been carried out, studies that have been conducted on agreement attraction in Turkish have been very limited. In fact, @Lago is currently the only study that explores this phenomenon in Turkish. @Lago  makes use of genitive-possessive structures in the subject position, in which the possessive-marked noun is the head of the noun phrase which acts as the grammatical agreement controller, and the genitive noun serves as a potential attractor. In a speeded acceptability judgment study, @Lago found a significant effect of number agreement attraction. However, the interpretation of their findings may be a result of the fact that non-subjecthood cues originate from their use of morphologically ambiguous forms of the possessive. In the possessive forms that are used, all the head nouns in their stimuli are ambigiuous between possessive and accusative.

@@

df_lagoetal_unp <- df_lagoetal %>% 
                    subset(is.na(Grammatical)) %>%
                    dplyr::select(-Grammatical:-Label)
df_lagoetal_attr <- df_lagoetal %>% 
                  subset(!is.na(Grammatical)) %>%
                  dplyr::select(-Distance:-NewCond)
```

@@

```{r loadFrequencies, eval =T, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}

word_freq <- readxl::read_excel("../Data/freq.xlsx", sheet = 1)
word_freq$freq_percentage %<>% as.numeric()
word_freq %<>% dplyr::select(-freq_standardized, -freq_percentage, -word)
word_freq %<>% tidyr::spread(place, freq_count)

word_freq %<>% dplyr::group_by(exp) %>% 
               dplyr::mutate( cat_n1 = ifelse(n1 > median(n1), "high", "low"),
                              cat_n2 = ifelse(n2 > median(n2), "high", "low"),
                              freqlog_n1 = scale(log(n1)),
                              freqlog_n2 = scale(log(n2)),
                              freqlog_n1n2 = log(n1) - log(n2)
                            ) %>% 
                ungroup()
word_freq$cat_n1 %<>% as.factor()
word_freq$cat_n2 %<>% as.factor()

word_freq_exp1 <- word_freq %>% 
                    subset(exp == "up") %>% 
                    dplyr::select(-exp)
word_freq_lago <- word_freq %>% 
                    subset(exp == "lago") %>% 
                    dplyr::select(-exp)

@@

df_merge_exp1 <- data_clean %>% ungroup() %>% 
                      dplyr::select(source=experiment, 
                                    grammatical, attractor_num,
                                    # condition,
                                    subject, item=Item,
                                    ResponseYes, RT)
df_merge_exp1$experiment <- "Experiment 1"
df_merge_exp1$grammatical <- with(df_merge_exp1, 
                                  ifelse(grammatical == "gram", 
                                          "grammatical",
                                          "ungrammatical"))
df_merge_exp1$attractor_num <- with(df_merge_exp1, 
                                    ifelse(attractor_num == "pl",
                                           "plural", 
                                           "singular"))
df_merge_exp1$item %<>% as.integer()
df_merge_exp1$subject %<>% as.character()
df_merge_exp1 %<>% left_join(word_freq_exp1, by = "item")

df_lagoetal_attr %<>% mutate(ResponseYes = (Response == "yes") )
df_merge_lago <- df_lagoetal_attr %>% ungroup() %>% 
                      dplyr::select(grammatical=Grammatical,
                                    attractor_num=Attractor,
                                    #condition=Condition, 
                                    subject=Participant, 
                                    item=Item,
                                    ResponseYes, RT)
df_merge_lago %<>% left_join(word_freq_lago, by = "item")
df_merge_lago$experiment <- "Lago et al. (2018)" 
df_merge_lago$source <- NA
df_merge_lago$item %<>% add(1000)

@@

avg_clean <- 
  df_merged %>%
     group_by(experiment, source, grammatical, attractor_num) %>%
     summarize(avgRT = mean(RT), p_yes = mean(ResponseYes, na.rm = T))

# ##
# avg_clean1 <- df_merged %>%
#                 group_by(experiment, source, grammatical, attractor_num, cat_n1) %>%
#                 summarize(avgRT = mean(RT), p_yes = mean(ResponseYes, na.rm = T))
# 
# p_avg_resp1 <- ggplot(avg_clean1 %>% subset(is.na(source) | source != "filler"), 
#                      aes(grammatical, p_yes, #linetype = attractor_num, 
#                          color = attractor_num, linetype = cat_n1, 
#                          group = paste(attractor_num,cat_n1) )) + 
#                 geom_point() + geom_line() + 
#                 facet_wrap(cat_n1~experiment)
# p_avg_resp1
# 
# ##
# ##
# 
# avg_clean2 <- df_merged %>%
#                 group_by(experiment, source, grammatical, attractor_num, cat_n2) %>%
#                 summarize(avgRT = mean(RT), p_yes = mean(ResponseYes, na.rm = T))
# 
# p_avg_resp2 <- ggplot(avg_clean2 %>% subset(is.na(source) | source != "filler"), 
#                      aes(grammatical, p_yes, #linetype = attractor_num, 
#                          color = attractor_num, linetype = cat_n2, 
#                          group = paste(attractor_num,cat_n2) )) + 
#                 geom_point() + geom_line() + 
#                 facet_wrap(cat_n2~experiment)
# p_avg_resp2
# 
# ##
# ##
# 
# avg_clean3 <- df_merged %>%
#                 group_by(experiment, source, grammatical, attractor_num, cat_n1n2) %>%
#                 summarize(avgRT = mean(RT), p_yes = mean(ResponseYes, na.rm = T))
# 
# p_avg_resp3 <- ggplot(avg_clean3 %>% subset(is.na(source) | source != "filler"), 
#                      aes(grammatical, p_yes, #linetype = attractor_num, 
#                          color = attractor_num, linetype = cat_n1n2, 
#                          group = paste(attractor_num,cat_n1n2) )) + 
#                 geom_point() + geom_line() + 
#                 facet_wrap(~experiment)
# p_avg_resp3
# 
# ##

@@

```{r exp1ModelResponse, eval = T}

@@

df_merged %<>% within(., {
  cGrammatical <- ifelse(grammatical == "grammatical", .5, -.5)
  cUngrammatical <- ifelse(grammatical == "ungrammatical", .5, -.5)
  cAttractorPlural <- ifelse(attractor_num == "plural", .5, -.5)
  cEndsInVowel <- ifelse(experiment == "Experiment 1", .5, -.5)
  cFreqlog_n1n2 <- scale(freqlog_n1n2)
  cFreqlog_n1 <- scale(freqlog_n1)
  cFreqlog_n2 <- scale(freqlog_n2)
})
df_merged_nofillers <- df_merged %>% subset(is.na(source) | source != "filler")

# ## test model parameterization using a simple GLM first
# m <- glm(ResponseYes ~ (cFreqlog_n1 + cFreqlog_n2) *
#            cEndsInVowel * cUngrammatical * cAttractorPlural,
#          data = df_merged_nofillers,
#          family = binomial("probit"))
# summary(m)

@@

#with(df_merged_nofillers, cor(cFreqlog_n1, cFreqlog_n2))
#with(df_merged_nofillers, cor(cFreqlog_n1, cFreqlog_n1n2))
#with(df_merged_nofillers, cor(cFreqlog_n2, cFreqlog_n1n2))
#
# m <- glm(ResponseYes ~ (cFreqlog_n1n2) * #+ cFreqlog_n1n2
#                         cEndsInVowel * cUngrammatical * cAttractorPlural,
#                    data = df_merged_nofillers,
#                    family = "binomial")
# summary(m)

# + cFreqlog_n1n2
m_responses <- brm(ResponseYes ~ (cFreqlog_n1 + cFreqlog_n2) *
                                  cEndsInVowel * cUngrammatical * cAttractorPlural + 
                                 ((cFreqlog_n1 + cFreqlog_n2) * cUngrammatical * cAttractorPlural + 1| subject) +
                                 (cUngrammatical * cAttractorPlural + 1| item),
                   data = df_merged_nofillers,
                   init = "0",
                   family = bernoulli("probit"),
                   file = fname_exp1_responses, 
                   chains = n_chains, cores = n_cores, iter = n_iter)

# + 
m_responses2 <- brm(ResponseYes ~ (cFreqlog_n1n2) *
                                  cEndsInVowel * cUngrammatical * cAttractorPlural + 
                                 ((cFreqlog_n1n2) * cUngrammatical * cAttractorPlural + 1| subject) +
                                 (cUngrammatical * cAttractorPlural + 1| item),
                   data = df_merged_nofillers,
                   init = "0",
                   family = bernoulli("probit"),
                   file = paste0(fname_exp1_responses, "2"), 
                   chains = n_chains, cores = n_cores, iter = n_iter)

@@

```{r exp1ResponseModelPlot, eval = T, fig.height= 2, message=FALSE, warning=FALSE, fig.cap="Estimates and 95% credible intervals for the analysis of the probability of a 'yes' response."}

contrast_names <- c("cUngrammatical" = "Ungrammaticality",
                    "cAttractorPlural" = "Plural Attactor",
                    "cEndsInVowel"="Vowel-ending",
                    "cUngrammatical:cAttractorPlural" = "Ungrammaticality * Plural Attactor",
                    "cEndsInVowel:cUngrammatical"="Vowel-ending * Ungrammaticality",
                    "cEndsInVowel:cAttractorPlural"="Vowel-ending * Plural Attactor",
                    "cEndsInVowel:cUngrammatical:cAttractorPlural"="Vowel-ending * Ungrammaticality * Plural Attactor"
                    )

### 2019-07-03 Paper/AgreementAttraction.tex
Some stuff. I don't remember.

\begin{figure}
\centering
\includegraphics{AgreementAttraction_files/figure-latex/exp1AvgResponseLagoEtAl-1.pdf}
\caption{\label{fig:exp1AvgResponseLagoEtAl}Estimates and 95\% credible
intervals for the analysis of the probability of a `yes' response.}
\end{figure}

>>>>

\begin{figure}
\centering
\includegraphics{AgreementAttraction_files/figure-latex/exp1ResponseModelPlot-1.pdf}
\caption{\label{fig:exp1ResponseModelPlot}Estimates and 95\% credible
intervals for the analysis of the probability of a `yes' response.}
\end{figure}
